---
title: "Analysis of Circumcision Data"
author: "Ramon + Sean"
date: today
format:
  html:
    toc: true
    page-layout: full
    embed-resources: true
    code-fold: true
    code-tools: true
    link-external-newwindow: true
execute:
  warning: false
from: markdown+emoji 
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: loading-packages-data

library(tidyverse)
library(DT)
```


# Overview + Highlights

This analysis analyzes []

In this sort of analysis, our expectation would be that shaft samples look the same between circumcised and uncircumcised individuals, and that the glans samples react to virus differently and/or have different resident immune cell populations.

- There were x,y,z differences that we found

- These differences were/were not the same in shaft/glans

- We did not detect differences between glans/shaft or circumcised/uncircumcised for the distance that virions penetrated


## Bigger Plots


## Specific Insights



## Remaining Questions/Concerns


## Conclusions + Future Analysis
- A lot of the emmeans models perform kinda poorly, and there isn't much we can do about this I feel like


# In Depth Analysis

## Virion Particle count


## Mast Cell Count Analysis


## Immunofluorescence Cell Distribution

We also had data from an immunofluorescence experiment, in which these samples of glans/shaft had been stained with CD3, CD4, and CCR10. Approximate cell types based on these markers are delineated below:

```{r}
#| label: cell-type-tibble

tribble(
  ~ "IF Expression", ~ "Expected Cell Type",
  "CD3+CD4+CCR10+", "Th22 (Tissue Homing T cell)",
  "CD3+CCR10+", "CD8+ Cutaneous Lymphoid Antigen T cell",
  "CD4+CCR10+", "Langerhans skin cell",
  "CD3+CD4+", "CD4+ T cell (likely Th1, Th2, Th9, or Th17)",
  "CD3+", "Other T cell",
  "CD4+", "HIV susceptible innate immune cells (macrophages, mast cells, etc)",
  "CCR10+", "Melanocytes"
) |>
  gt::gt() |>
  gt::tab_header(title = "Expected IF Cell Types")
```

Here, looking at the number of cells per sample isn't the most interesting, as this is likely just a result of how the tissue was cut. What *would* be interesting would be to look at if the ***distribution***, i.e. how far any particular cell type was from the beginning of the epithelial layer, of these cells changed as a result of the virion infection. 

[INSERT ANALYSIS]


### Adding Back virion penetration data

We can also then go back to correlate this with virion penetration data:

```{r}
#| label: correlation-plot-one

load(file = "figures/plots/correlation_plot_1.rda")

correlation_plot_1
```







